---
title: "Data622-HW-1"
author: "Deepak Mongia"
date: "3/30/2020"
output: html_document
---

You just recently joined a datascience team.

There are two datasets junk1.txt and junk2.csv
They have two options
1. They can go back to the client and ask for more data
to remedy problems with the data.
2. They can accept the data and undertake a major analytics exercise.

The team is relying on your dsc skills to determine how they
should proceed.

Can you explore the data and recommend actions for each file
enumerating the reasons.

```{r}
library(ggplot2)
library(cowplot)

junk1_ds <- read.table("https://raw.githubusercontent.com/deepakmongia/Data622/master/Homework-1/Data/junk1.txt",header = TRUE)

head(junk1_ds)
summary(junk1_ds)

```

Let us look at the first file - Junk1

```{r}
unique(junk1_ds$class)

junk1_ds$class = as.factor(junk1_ds$class)

summary(junk1_ds)

str(junk1_ds)

apply(junk1_ds[,1:2], 2, sd)
```

Plotting the bixplots to see if there are any outliers:
```{r}
boxplot(junk1_ds[,1:2])
```
As we see above, there are no outliers.
Verifying that there are no outliers:
```{r}
OutVals_a_1 = boxplot(junk1_ds$a, plot=FALSE)$out
print(OutVals_a_1)

OutVals_b_1 = boxplot(junk1_ds$b, plot=FALSE)$out
print(OutVals_b_1)
```

So, it is confimed that there are no outliers in the data.

Also from the table of the classes, the number of observations are equal for both the classes.

So, we don't see any issues with the data - junk1, and it can be used with the analytics exercise without any further changes.


Now, let us import the 2nd file - Junk2, and do some exploratory data analysis on the same.
```{r}
junk2_ds <- read.csv("https://raw.githubusercontent.com/deepakmongia/Data622/master/Homework-1/Data/junk2.csv",
                     header = TRUE)

```

Doing some basic EDA
```{r}
dim(junk2_ds)

head(junk2_ds)

summary(junk2_ds)
```

Converting class feature into a factor
```{r}
junk2_ds$class <- as.factor(junk2_ds$class)

table(junk2_ds$class)
```

So, the dataset is imbalanced as the class '1' is having only 250 observations out of 4000 total.

Plotting the boxplots to see the data ranges and see if there are any outliers.

```{r}
boxplot(junk2_ds[,1:2])
```

Looks like there are outliers in both the independent features - a and b.
Checking the outlier values.
```{r}
OutVals_a_2 = boxplot(junk2_ds$a, plot=FALSE)$out
print(OutVals_a_2)

OutVals_b_2 = boxplot(junk2_ds$b, plot=FALSE)$out
print(OutVals_b_2)
```

So, there are outliers in the data.

For junk2 dataset:
1) there are outliers in the data
2) the data is highly imbalanced

So we should go back to the source team to get more samples of the class with the lower number of observations which is class '1'.